---
title: "<u>Homework 5 - Matrices, Lists, and Data frames</u>"
author: "Grace (Rei) Jia"
date: "2025-02-12"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<u>Question 1:</u> Assign to the variable n_dims a single random integer between 3 and 10.
```{r}
n_dims <- sample(seq(3:10), size=1)
```

Create a vector of consecutive integers from 1 to n_dims^2
```{r}
vector <- seq(1:n_dims^2)
```

Use the sample function to randomly reshuffle these values. Create a square matrix with these elements. Print out the matrix.
```{r}
matrix1 <- matrix(sample(vector), nrow=n_dims)
```

Find a function in r to transpose the matrix. Print it out again and note how it has changed.
```{r}
print(t(matrix1))
```

Calculate the sum and the mean of the elements in the first row and then the last row.
```{r}
mean(matrix1[1,])
sum(matrix1[1,])

mean(matrix1[2,])
sum(matrix1[2,])
```

Read about the eigen() function and use it on your matrix. Look carefully at the elements of $values and $vectors in the output. What kind of numbers are these?
```{r}
matrix_eigen <- eigen(matrix1)
print(matrix_eigen)
```

Dig in with the typeof() function to figure out their type.
```{r}
typeof(matrix_eigen$values)
typeof(matrix_eigen$vectors)
```

<u>Question 2:</u> Create a list with the following named elements:

my_matrix, which is a 4 x 4 matrix filled with random uniform values
```{r}
my_matrix <- matrix(data=runif(16), nrow=4)
print(my_matrix)
```

my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
```{r}
my_logical <- (runif(100) > .8)
print(my_logical)
```

my_letters, which is a 26-element vector of all the lower-case letters in random order.
```{r}
my_letters <- sample(letters[1:26])
print(my_letters)
```
Create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.

```{r}
list1 <- list(my_matrix, my_logical, my_letters)

new_list <- list(my_matrix[2,2], my_logical[2], my_letters[2])
```

Use the typeof() function to confirm the underlying data types of each component in this list

```{r}
typeof(new_list[[1]])
typeof(new_list[[2]])
typeof(new_list[[3]])
```

Combine the underlying elements from the new list into a single atomic vector with the c() function. What is the data type of this vector?

```{r}
combined_elements_vector <- c(new_list[[1]], new_list[[2]], new_list[[3]])
print(combined_elements_vector)
typeof(combined_elements_vector)
```

<u>Question 3:</u> Create a data frame with the two variables (= columns) and 26 cases (= rows) below. Call the first variable my_unis and fill it with 26 random uniform values from 0 to 10. Call the second variable my_letters and fill it with 26 capital letters in random order.
```{r}
dataframe <- data.frame(my_unis=runif(26, min=0, max=10), my_letters=sample(LETTERS[1:26]))
```

For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
```{r}
dataframe[sample(dataframe$my_unis, size=4, replace=TRUE),1]=NA
```

For the first variable, write a single line of R code to identify which rows have the missing values.
```{r}
print(which(is.na(dataframe$my_unis)))
```
Re-order the entire data frame to arrange the second variable in alphabetical order
```{r}
dataframe[order(dataframe$my_letters),2]
```
Calculate the column mean for the first variable.
```{r}
mean(dataframe$my_unis, na.rm=TRUE)
```
