---
title: "<u>Homework 6 – Creating Fake Data Sets To Explore Hypotheses</u>"
author: "Grace (Rei) Jia"
date: "2025-02-19"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My dataset was made at random for this homework assignment, very loosely based on biodiversity of alpine plant populations in the northeast. 

Making data and a quick summary of each plant population:
```{r}
plantpopulation1 <- rnorm(n=10, mean = 150, sd = 20)
plantpopulation2 <- rnorm(n=10, mean = 200, sd = 50)
summary(plantpopulation1)
summary(plantpopulation2)
```

Putting it into a data frame:
```{r}
Plant_Populations <- c(plantpopulation1, plantpopulation2)
Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
plant_df <- data.frame(Plant_Populations, Group_Name)
print(plant_df)
```

Plotting the data
```{r}
boxplot(Plant_Populations~Group_Name,data=plant_df, main="Alpine Plant Population Density",
        xlab="Plant Groups", ylab="Population Size")
```

```{r}
hist(plant_df$Plant_Populations, 
     main = paste("Histogram of Plant Populations 1 and 2"),
     xlab = "Population Density", ylab = "Frequency")
```

Using the same parameters as in the previous step, I am now using a for loop to simulate 20 different iterations of this combined plant populations. My original sample size was 10.

```{r}
n_iterations <- 1:20

for (i in n_iterations){
  plantpopulation1 <- rnorm(n=10, mean = 150, sd = 20)
  plantpopulation2 <- rnorm(n=10, mean = 200, sd = 50)
  Plant_Populations <- c(plantpopulation1, plantpopulation2)
  Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
  plant_df <- data.frame(Plant_Populations, Group_Name)
  test_result <- t.test(Plant_Populations~Group_Name, data = plant_df)
  n_iterations[i] <- test_result$p.value
  cat("Iteration=",i,"P-value=")
print(test_result$p.value)
}
```
From the 20 iterations above, I determined that several have a p-value of < 0.05 (which changes every time I run the for loop).
```{r}
which(n_iterations < 0.05)
```

Here I am running the same for loop but I changed the sample size to 5.

```{r}
n_iterations <- 1:20

for (i in n_iterations){
  plantpopulation1 <- rnorm(n=5, mean = 150, sd = 20)
  plantpopulation2 <- rnorm(n=5, mean = 200, sd = 50)
  Plant_Populations <- c(plantpopulation1, plantpopulation2)
  Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
  plant_df <- data.frame(Plant_Populations, Group_Name)
  test_result <- t.test(Plant_Populations~Group_Name, data = plant_df)
  n_iterations[i] <- test_result$p.value
  cat("Iteration=",i,"P-value=")
print(test_result$p.value)
}
```

Again, from the 20 iterations above with a sample size of 5, I determined that several have a p-value of < 0.05 (which changes every time I run the for loop).
```{r}
which(n_iterations < 0.05)
```

Continuing to use this for loop, I have determined that the smallest sample size that is still statistically significant is 2, which is rather shocking. Evidence below!

```{r}
n_iterations <- 1:20

for (i in n_iterations){
  plantpopulation1 <- rnorm(n=2, mean = 150, sd = 20)
  plantpopulation2 <- rnorm(n=2, mean = 200, sd = 50)
  Plant_Populations <- c(plantpopulation1, plantpopulation2)
  Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
  plant_df <- data.frame(Plant_Populations, Group_Name)
  test_result <- t.test(Plant_Populations~Group_Name, data = plant_df)
  n_iterations[i] <- test_result$p.value
  cat("Iteration=",i,"P-value=")
print(test_result$p.value)
}
```
```{r}
which(n_iterations < 0.05)
```
