---
title: "<u>Homework 6 – Creating Fake Data Sets To Explore Hypotheses</u>"
author: "Grace (Rei) Jia"
date: "2025-02-19"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My dataset was made at random for this homework assignment, very loosely based on biodiversity of alpine plant populations in the northeast. 

Making data and a quick summary of each plant population:
```{r}
plantpopulation1 <- rnorm(n=10, mean = 57, sd = 13)
plantpopulation2 <- rnorm(n=10, mean = 29, sd = 20)
summary(plantpopulation1)
summary(plantpopulation2)
```

Putting it into a data frame:
```{r}
Plant_Populations <- c(plantpopulation1, plantpopulation2)
Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
plant_df <- data.frame(Plant_Populations, Group_Name)
print(plant_df)
```

Analyzing the data by obtaining the mean and standard deviations from both plantpopulation1 and plantpopulation2.
```{r}
mean(plant_df$Plant_Populations)
sd(plant_df$Plant_Populations)

```

Plotting the data
```{r}
boxplot(Plant_Populations~Group_Name,data=plant_df, main="Alpine Plant Population Density",
        xlab="Plant Groups", ylab="Population Size")
```

```{r}
hist(plant_df$Plant_Populations, 
     main = paste("Histogram of Plant Populations 1 and 2"),
     xlab = "Population Density", ylab = "Frequency")
```

```{r}
n_iterations <- 1:10

for (i in n_iterations) {
  plantpopulation1 <- rnorm(n=i, mean = 57, sd = 13)
  plantpopulation2 <- rnorm(n=i, mean = 29, sd = 20)
  Plant_Populations <- c(plantpopulation1, plantpopulation2)
  Group_Name <- c(rep("Plant_Population_1", length(plantpopulation1)),rep("Plant_Population_2",length(plantpopulation2)))
  test_result <- t.test(Plant_Populations~Group_Name, data = plant_df)
  cat("Iteration=",i,"P-value=")
print(test_result$p.value)
}
```
- Not sure why, but all of my values are the same

