---
title: "<u>Homework 4 - Atomic Vectors and Basic R</u>"
author: "Grace (Rei) Jia"
date: "2025-02-05"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<u>Question 1:</u> Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.
```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```
a)
```{r}
z <- x^(a^b)
print(z)
```

b)
```{r}
z <- (x^a)^b
print(z)
```

c)
```{r}
z <- 3*x^3 + 2*x^2 + 1
print(z)

```

<u>Question 2:</u>Using the rep and seq functions, create the following vectors:

a) (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)
```{r}
c(seq(1,8),seq(7,1))
```

b) (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)
```{r}
rep(x=seq(1,5),times=seq(1,5))
```

c) (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r}
rep(x=seq(5,1),times=seq(1,5))
```

<u>Question 3:</u> Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.
```{r}
cartesian_coords <- runif(2)
print(cartesian_coords)
```
```{r}
r <- sqrt(cartesian_coords[1])^2 + (cartesian_coords[2])^2
theta <- atan(cartesian_coords[2]/cartesian_coords[1])
polar_coords <- c(r, theta)
print(polar_coords)
```

<u>Question 4:</u> Create a vector queue <- c("sheep", "fox", "owl", "ant") where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update queue as:
```{r}
queue <- c("sheep", "fox", "owl", "ant")
```

a) the serpent arrives and gets in line;
```{r}
queue <- c(queue, "serpent")
print(queue)
```

b) the sheep enters the ark;
```{r}
queue <- queue[-1]
print(queue)
```

c) the donkey arrives and talks his way to the front of the line;
```{r}
queue <- c("donkey", queue)
print(queue)
```

d) the serpent gets impatient and leaves;
```{r}
queue <- queue[-5]
print(queue)
```

e) the owl gets bored and leaves;
```{r}
queue <- queue[-3]
print(queue)
```

f) the aphid arrives and the ant invites him to cut in line.
```{r}
queue <- c(queue[1:2], "aphid", queue[3])
print(queue)
```
g) Finally, determine the position of the aphid in the line.
```{r}
print(length(queue[-1]))
```

<u>Question 5:</u> Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7. 
```{r}
q5vector <- seq(1,100)
print(which(q5vector%%2 & q5vector%%3 & q5vector%%7))
```
