---
title: "Homework 10 – Advanced ggplotting"
author: "Grace (Rei) Jia"
date: "2025-04-09"
output: 
  html_document:
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(ggridges)
library(ggbeeswarm)
library(GGally)
library(ggpie)
library(ggmosaic)
library(scatterpie)
library(waffle)
library(DescTools)
library(treemap)
library(devtools)
library(extrafont)
```

The following code for Steps 1 and 2 were completed using code provided by TidyTuesday and can be found [here](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-03-18/readme.md)

Reading in Palm Trees dataset from [TidyTuesday](https://github.com/rfordatascience/tidytuesday)
```{r message=FALSE, warning=FALSE}
palmtrees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-03-18/palmtrees.csv',show_col_types = FALSE)
```

Looking at the data frame.
```{r message=FALSE, warning=FALSE}
head(palmtrees)
```
<u>Barplot:</u>
```{r message=FALSE, warning=FALSE}
palm_barplot <- ggplot(palmtrees) + 
                          aes(x=palm_subfamily,fill=fruit_shape) +
                          geom_bar(position="dodge",color="black",size=0.5) +
                          labs(x = "Subfamily", 
                               y = " # of Species within Subfamily", 
                               fill = "Fruit Shape") +
                          scale_fill_brewer(palette = "Pastel1") +
                          theme(text = element_text(family = "Helvetica"),
                                  axis.title.x = element_text(size = 12),
                                  axis.text.x = element_text(size = 10),
                                  axis.title.y = element_text(size = 10))
                          
palm_barplot
```

<u>Pie chart:</u>
<br>- In this pie chart, I am looking at the frequency of each palm tree subfamilies within the data set. It shows that the subfamily, Arecoideae, is most common.</br>
```{r message=FALSE, warning=FALSE}
palm_piechart <- ggpie::ggpie(data=palmtrees,
                             group_key="palm_subfamily",
                             count_type="full",
                             label_info="ratio", 
                             label_type="none") +
                            scale_fill_brewer(palette = "Pastel1") +
                            labs(fill = "Subfamily Name") +
                            theme(text = element_text(family = "Helvetica"))
        
palm_piechart
```

For the remaining figures, I subset the data set by subfamily using the following code:
```{r message=FALSE, warning=FALSE}
arecoideae <- palmtrees[which(palmtrees$palm_subfamily == "Arecoideae"),]
calamoideae <- palmtrees[which(palmtrees$palm_subfamily == "Calamoideae"),]
ceroxyloideae <- palmtrees[which(palmtrees$palm_subfamily == "Ceroxyloideae"),]
coryphoideae <- palmtrees[which(palmtrees$palm_subfamily == "Coryphoideae"),]
nypoideae <- palmtrees[which(palmtrees$palm_subfamily == "Nypoideae"),]
```


<u>Beeswarm Plot:</u> 
<br>- Using a beeswarm plot, I wanted to see if there is a relationship between palm trees in the subfamily, Calamoideae, and whether species within that subfamily exhibit "climbing" behavior (climb, don't climb, or do both) in relation to their maximum petiole length. It shows that species that climb have a longer petiole length compared to non-climbers and palm trees in both categories.</br>
```{r message=FALSE, warning=FALSE}
palm_beeswarm <- ggplot(data=calamoideae) +
                            aes(x=climbing,y=max__petiole_length_m,color=climbing) + 
                            ggbeeswarm::geom_beeswarm(method = "center",size=3) +
                            scale_color_brewer(palette = "Accent") +
                            labs(x = "Climbing Type", y = "Maximum Petiole Length") +
                            theme(legend.position="none") +
                            theme(text = element_text(family = "Helvetica"),
                                  axis.title.x = element_text(size = 12),
                                  axis.text.x = element_text(size = 10),
                                  axis.title.y = element_text(size = 12))

palm_beeswarm
```

<u>Waffle plot:</u> 
<br>- Using a waffle plot, I wanted to see how many tribes are within the subfamily, Coryphoideae. This waffle plot shows the number of species within each tribe.</br>
```{r message=FALSE, warning=FALSE}
tabled_data <- as.data.frame(table(class=coryphoideae$palm_tribe))

palm_waffle <- ggplot(data=tabled_data) +
                          aes(fill = class, values = Freq) +
                          waffle::geom_waffle(n_rows = 10, size = 0.3, colour = "white") +
                          scale_fill_manual(name = "Tribe Names:",
                                            values = c("#698B69", "#528B8B","#CD8C95", "#B2DFEE", "#EE9572","#CDCDC1","#CD96CD","#CDBA96"),
                                            labels = c("Borasseae", "Caryoteae", "Chuniophoeniceae","Corypheae","Cryosophileae","Phoeniceae","Sabaleae","Trachycarpeae")) +
                          coord_equal() +
                          theme_void() +
                          theme(legend.position = "bottom", text = element_text(family = "Helvetica"))

palm_waffle
```

<u>Ridgeline plot:</u>
<br>- This ridgeline plot shows the distribution of maximum stem height (m) by subfamily.</br>
```{r message=FALSE, warning=FALSE}
palm_subfamily_df <- data.frame(palmtrees$palm_subfamily)
palm_stemheight_df <- data.frame(palmtrees$max_stem_height_m)
palm_df <- cbind(palm_subfamily_df, palm_stemheight_df)
palm_df_clean <- palm_df[complete.cases(palm_df),]

palm_ridgeline <- ggplot(data=palm_df_clean,
                         aes(x=palmtrees.max_stem_height_m,y=palmtrees.palm_subfamily, fill = palmtrees.palm_subfamily)) + 
                         scale_fill_brewer(palette = "Accent") +
                         ggridges::geom_density_ridges(scale = 0.9) +
                         ggridges::theme_ridges() +
                         scale_x_continuous(limits = c(-10,100)) +
                         xlab("Maximum stem height (m)") +
                         ylab("Subfamily") +
                         theme(legend.position = "none", text = element_text(family = "Helvetica"))

palm_ridgeline
```

<u>Saving each plots:</u>
```{r message=FALSE, warning=FALSE}
ggsave(palm_piechart,file="palm_piechart.png",width=10,height=6,bg = "white")
ggsave(palm_beeswarm,file="palm_beeswarm.png",width=10,height=6,bg = "white")
ggsave(palm_waffle,file="palm_waffle.png",width=10,height=6,bg = "white")
ggsave(palm_ridgeline,file="palm_ridgeline.png",width=10,height=6,bg = "white")
```
