---
title: "<u>Homework 8 – Simulating and Fitting Data Distributions</u>"
author: "Grace (Rei) Jia"
date: "2025-02-26"
output: 
  html_document:
    theme: journal
---
<break></break>

Data for this homework was obtained from www.data.gov and is titled "Harmful Algal Bloom Statewide Occurrence Summary: 2012-2018" from the State of New York. For context, HABs/habs = Harmful Algal Blooms

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Importing necessary libraries:
```{r}
library(ggplot2)
library(MASS)
```
Importing dataset:
```{r}
habs <- read.table("HABs_2012-2018.csv", header=TRUE,sep=",")
```

Cleaning up data and assigning variable of interest:
```{r}
habs <- na.omit(habs)
habs$myVar <- habs$Number.of.Weeks.on.DEC.Notification.List
```

Summarizing dataset:
```{r}
str(habs$myVar)
summary(habs$myVar)
```

Histogram of Number of Weeks HABs were listed on NYSDEC Notification List:
```{r}
p1 <- ggplot(data=habs, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```

Adding empirical density curve:
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```

Maximum likelihood parameters for normal distribution
```{r}
normPars <- fitdistr(habs$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```
Plotting normal probability density on the histogram:
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(habs$myVar),len=length(habs$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(habs$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

Plotting exponential probability density on the histogram:
```{r}
expoPars <- fitdistr(habs$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(habs$myVar), args = list(rate=rateML))
 p1 + stat + stat2
```



Plotting uniform probability density on the histogram:
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(habs$myVar), args = list(min=min(habs$myVar), max=max(habs$myVar)))
 p1 + stat + stat2 + stat3
```

Plotting gamma probability density on the histogram:
```{r}
gammaPars <- fitdistr(habs$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(habs$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```



Plotting beta probability density on the histogram:
```{r}
pSpecial <- ggplot(data=habs, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=habs$myVar/max(habs$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(habs$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```


<u>Simulated New Dataset</u>

Creating a simulated dataset based on gamma probability density. Here I used "rgamma" instead of "dgamma", which will generate random values.
```{r}
simulated_dataset <- rgamma(n=length(habs$myVar),shape=shapeML,rate=rateML)
simulated_z <- 1:length(habs$myVar)
simulated_df <- data.frame(simulated_dataset,simulated_z)
```

Summary of simulated dataset:
```{r}
str(simulated_dataset)
summary(simulated_dataset)
```

Plotting the histogram based on gamma probability density, which appears to best fit the data compared to normal, exponential, uniform, and beta probability.
```{r}
p2 <- ggplot(data=simulated_df, aes(x=simulated_dataset, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",linewidth=0.2) 
p2 + stat4
```
